Uncertainty Quantification and Inference of Manning's Friction Coefficient using DART Buoys Data during \tohoku Tsunami

### 2014-01-20 paper/mcmc.tex
new jpdfs and discussion of correlations

Next, the computed MCMC chains are used to determine the posterior 
distributions using kernel density estimation (KDE)
~\citep{Parzen1962,Silverman1986}.  The resulting posterior pdfs 
of the three Manning's $n$ coefficients $n_1,n_2,n_3$ are shown 
in Figure~\ref{fig:pdfs} in addition to the pdfs of $\sigma^2$. 
Note that the first 10000 iterates, associated with the burn-in period, were discarded.  
As expected from the chains shown in Figure~\ref{fig:mcmc}, the posterior pdfs of $n_1$ appear to be fairly flat, 
and similar to the uniform prior; an indication that 
the observed data were not useful to refine our prior knowledge for $n_1$.,
In contrast, the posterior pdf of $n_2$ exhibits well-defined peak, 
with a Maximum A Posteriori (MAP) estimate of around 0.011
and an extended tail towards the larger Manning's $n$ coefficient values.
For $n_3$, we observe a posterior that has a well defined peak
of 0.18 but no clear pdf shape. The posterior distribution of the variance 
appears to be well-defined of Gaussian-like shape. 
The MAP value of $\sigma^2=0.021$ can be used to estimate the water surface elevation standard 
deviation estimated to be 0.145~$m$. 
This value is a reflection of the mismatch between the model and 
observed data. \comment{marginal "", correlation ''', joint}.

>>>>

Next, the computed MCMC chains are used to determine the marginalized posterior 
distributions using kernel density estimation (KDE)~\citep{Parzen1962,Silverman1986}.  
The resulting marginalized posterior pdfs of the three Manning's $n$ coefficients 
$n_1,n_2,n_3$ are shown in Figure~\ref{fig:pdfs} in addition to the pdfs of $\sigma^2$. 
Note that the first 10000 iterates, associated with the burn-in period, were discarded.  
As expected from the chains shown in Figure~\ref{fig:mcmc}, the marginalized posterior 
pdfs of $n_1$ appear to be fairly flat, and similar to the uniform prior; an indication 
that the observed data were not useful to refine our prior knowledge for $n_1$. In 
contrast, the posterior pdf of $n_2$ exhibits well-defined peak, with a Maximum A 
Posteriori (MAP) estimate of around 0.011 and an extended tail towards the larger 
Manning's $n$ coefficient values. For $n_3$, we observe a posterior that has a well 
defined peak of 0.18 but no clear pdf shape. The posterior distribution of the variance 
appears to be well-defined of Gaussian-like shape. The MAP value of $\sigma^2=0.021$ 
can be used to estimate the water surface elevation standard deviation estimated to 
be 0.145~$m$. This value is a reflection of the mismatch between the model and 
observed data.

We finally investigate the existence of any correlation arising between the inferred parameters.
This study is motivated by the fact that the Manning's $n$ coefficients have an important role 
in characterizing the tsunami model predictions. Moreover, it is instructive to analyze the 
correlation between the various $n$ parameters to gain some insight into their mutual dependence.
To this end, we compute the two-dimensional marginalized joint posteriors for each 
possible combination of the parameters $n_1$, $n_2$ and $n_3$, namely $p(n_1, n_2), p(n_1, n_3)$ and $p(n_2, n_3)$. 
Figure~\ref{fig:joint}(left) shows the scatter plots of the MCMC samples of the combinations of 
inferred parameters obtained from the chain samples of the marginalized posterior excluding 
the burn-in iterations. The plots reveal no clear correlation in all three cases. 
These samples can be readily exploited to reconstruct the joint density for each combination of parameters. 
The right column of Figure~\ref{fig:joint} shows the contour plots of the marginalized joint posteriors 
$p(n_1, n_2)$, $p(n_1, n_3)$ and $p(n_2, n_3)$. The contours are obtained from the
distribution reconstructed via KDE and shown for the cases corresponding to greater than $10\%$
of the maximum probability.

To quantify the degree of no-correlation between each possible pair of parameters, we compute 
the corresponding correlation coefficient which is a unitless scalar that falls in the range 
$(-1, 1)$, with $1$ indicating a perfect positive correlation, $−1$ a perfect negative correlation and $0$ for
a no-correlation case.
% $R(n_2, n_3)$, $R(n_2, \sigma^2)$, and $R(n_3, \sigma^2)$.
The correlation coefficient, $R(X, Y )$, between two random variables $X$ and $Y$ is defined as:

\begin{equation}
R(X,Y) = \frac{E[(X - \mu_X)(Y - \mu_Y )]}{\sigma_X \sigma_Y} 
\end{equation}
where $\sigma_X$ and  $\sigma_Y$ are the standard deviations of  $X$ and $Y$, $\mu_X$
and $\mu_Y$  are the means of  $X$ and $Y$ respectively;
and $E$ is the expectation operator.  In our case, the correlation coefficient
was less than $0.06$ for all possible combinations of parameters; again indicating no correlation
existing between the inferred parameters.
